### 2024-07-22 NN/tables/forward.tex
commit

    \begin{table*}
        \centering
        \begin{tabular}{lr}
\hline
 features                                                      &    R2 \\
\hline
 X\_D1                                                         & 15.01 \\
 flow\_rate2                                                   & 34.3  \\
 conc\_nano2                                                   & 10.78 \\
 Kfluid2                                                       & 49.29 \\
 heat\_flux2                                                   & -8.26 \\
 flow\_rate1,Kfluid2                                           & 52.94 \\
 conc\_nano1,Kfluid2                                           & 50.8  \\
 Kfluid1,Kfluid2                                               & 47.57 \\
 heat\_flux1,Kfluid2                                           & 52.31 \\
 X\_D1,Kfluid2                                                 & 62.32 \\
 flow\_rate2,Kfluid2                                           & 29.14 \\
 conc\_nano2,Kfluid2                                           & 61.05 \\
 heat\_flux2,Kfluid2                                           & 66.61 \\
 X\_D2,Kfluid2                                                 & 53.44 \\
 flow\_rate1,Kfluid2,heat\_flux2                               & 60.58 \\
 conc\_nano1,Kfluid2,heat\_flux2                               & 61.96 \\
 Kfluid1,Kfluid2,heat\_flux2                                   & 66.64 \\
 heat\_flux1,Kfluid2,heat\_flux2                               & 42.41 \\
 X\_D1,Kfluid2,heat\_flux2                                     & 66.55 \\
 flow\_rate2,Kfluid2,heat\_flux2                               & 55.24 \\
 conc\_nano2,Kfluid2,heat\_flux2                               & 67.09 \\
 X\_D2,Kfluid2,heat\_flux2                                     & 68.58 \\
 flow\_rate1,Kfluid2,heat\_flux2,X\_D2                         & 90.12 \\
 conc\_nano1,Kfluid2,heat\_flux2,X\_D2                         & 78.64 \\
 Kfluid1,Kfluid2,heat\_flux2,X\_D2                             & 88.23 \\
 heat\_flux1,Kfluid2,heat\_flux2,X\_D2                         & 49.47 \\
 X\_D1,Kfluid2,heat\_flux2,X\_D2                               & 79.9  \\
 flow\_rate2,Kfluid2,heat\_flux2,X\_D2                         & -0.88 \\
 conc\_nano2,Kfluid2,heat\_flux2,X\_D2                         & 63.17 \\
 conc\_nano1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1             & 95.97 \\
 Kfluid1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1                 & 76.54 \\
 heat\_flux1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1             & 60.12 \\
 X\_D1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1                   & 80.67 \\
 flow\_rate2,Kfluid2,heat\_flux2,X\_D2,flow\_rate1             & 68.87 \\
 conc\_nano2,Kfluid2,heat\_flux2,X\_D2,flow\_rate1             & 68.35 \\
 Kfluid1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1,conc\_nano1     & 80.62 \\
 heat\_flux1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1,conc\_nano1 & 67.03 \\
 X\_D1,Kfluid2,heat\_flux2,X\_D2,flow\_rate1,conc\_nano1       & 85.73 \\
 flow\_rate2,Kfluid2,heat\_flux2,X\_D2,flow\_rate1,conc\_nano1 & 75.7  \\
 conc\_nano2,Kfluid2,heat\_flux2,X\_D2,flow\_rate1,conc\_nano1 & 78.99 \\
\hline
\end{tabular}
        \caption{Results of Forward Feature Selection for different features}
        \label{forward}
    \end{table*}