### 2024-07-22 NN/tables/backward.tex
commit

    \begin{table*}
        \centering
        \begin{tabular}{lr}
\hline
 features                                                                                                     &    R2 \\
\hline
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2, X\_D2 & 82.69 \\
 conc\_nano1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2, X\_D2              & 62.45 \\
 flow\_rate1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2, X\_D2              & 81.63 \\
 flow\_rate1, conc\_nano1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2, X\_D2          & 73.64 \\
 flow\_rate1, conc\_nano1, Kfluid1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2, X\_D2              & 72.94 \\
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2, X\_D2        & 67.45 \\
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, X\_D1, conc\_nano2, Kfluid2, heat\_flux2, X\_D2              & 68.03 \\
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, Kfluid2, heat\_flux2, X\_D2              & 76.57 \\
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, heat\_flux2, X\_D2          & 38.28 \\
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, X\_D2              & 69.39 \\
 flow\_rate1, conc\_nano1, Kfluid1, heat\_flux1, X\_D1, flow\_rate2, conc\_nano2, Kfluid2, heat\_flux2        & 81.61 \\
\hline
\end{tabular}
        \caption{Results of Backward Feature Elimination}
        \label{backward}
    \end{table*}