### 2024-07-22 NN/tables/forward.tex

    \begin{table*}
        \centering
        \begin{tabular}{lr}
\hline
 features                                                             &    R2 \\
\hline
 X\_D1                                                                & 14.85 \\
 flow\_rate2                                                          & 27.69 \\
 conc\_nano2                                                          & 11.15 \\
 Kfluid2                                                              & 62.66 \\
 heat\_flux2                                                          & -4.92 \\
 flow\_rate1, Kfluid2                                                 & 75.33 \\
 conc\_nano1, Kfluid2                                                 & 64.49 \\
 Kfluid1, Kfluid2                                                     & 67.9  \\
 heat\_flux1, Kfluid2                                                 & 74.3  \\
 X\_D1, Kfluid2                                                       & 67.3  \\
 flow\_rate2, Kfluid2                                                 &  9.94 \\
 conc\_nano2, Kfluid2                                                 & 87.59 \\
 heat\_flux2, Kfluid2                                                 & 56.8  \\
 X\_D2, Kfluid2                                                       & 67.59 \\
 flow\_rate1, Kfluid2, conc\_nano2                                    & 60.42 \\
 conc\_nano1, Kfluid2, conc\_nano2                                    & 78.36 \\
 Kfluid1, Kfluid2, conc\_nano2                                        & 79.88 \\
 heat\_flux1, Kfluid2, conc\_nano2                                    & 90.37 \\
 X\_D1, Kfluid2, conc\_nano2                                          & 84.51 \\
 flow\_rate2, Kfluid2, conc\_nano2                                    & 65.14 \\
 heat\_flux2, Kfluid2, conc\_nano2                                    & 57.1  \\
 X\_D2, Kfluid2, conc\_nano2                                          & 54.6  \\
 flow\_rate1, Kfluid2, conc\_nano2, heat\_flux1                       & 62.62 \\
 conc\_nano1, Kfluid2, conc\_nano2, heat\_flux1                       & 92.87 \\
 Kfluid1, Kfluid2, conc\_nano2, heat\_flux1                           & 80.94 \\
 X\_D1, Kfluid2, conc\_nano2, heat\_flux1                             & 87.63 \\
 flow\_rate2, Kfluid2, conc\_nano2, heat\_flux1                       & 82.76 \\
 heat\_flux2, Kfluid2, conc\_nano2, heat\_flux1                       & 55.75 \\
 X\_D2, Kfluid2, conc\_nano2, heat\_flux1                             & 75.43 \\
 flow\_rate1, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1          & 49.56 \\
 Kfluid1, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1              & 93.06 \\
 X\_D1, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1                & 87.49 \\
 flow\_rate2, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1          & 69.35 \\
 heat\_flux2, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1          & 65.5  \\
 X\_D2, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1                & 64.18 \\
 flow\_rate1, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1, Kfluid1 & 84.19 \\
 X\_D1, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1, Kfluid1       & 87.54 \\
 flow\_rate2, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1, Kfluid1 & 76.61 \\
 heat\_flux2, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1, Kfluid1 & 71.81 \\
 X\_D2, Kfluid2, conc\_nano2, heat\_flux1, conc\_nano1, Kfluid1       & 77.34 \\
\hline
\end{tabular}
        \caption{Results of Forward Feature Selection for different features}
        \label{forward}
    \end{table*}