### 2024-08-08 NN/tables/forward.tex

    \begin{table*}[h]
        \centering
        \begin{tabular}{lr}
\hline
 features                                                         &     R2 \\
\hline
 pH Sol                                                           &  14.56 \\
 Hydro Time (h)                                                   &  26.4  \\
 Dip Time Sol (s)                                                 &  11.24 \\
 Clay Conc                                                        &  64.24 \\
 Cure Time (min)                                                  &   0.1  \\
 SG Coating, Clay Conc                                            &  64.93 \\
 Temp Alk Sol, Clay Conc                                          &  64.93 \\
 NaOH Wt\%, Clay Conc                                             &  64.93 \\
 Dip Time Alk (min), Clay Conc                                    &  79.07 \\
 pH Sol, Clay Conc                                                &  62.87 \\
 Hydro Time (h), Clay Conc                                        & -15.61 \\
 Dip Time Sol (s), Clay Conc                                      &  80.99 \\
 Cure Time (min), Clay Conc                                       &  74.1  \\
 Cure Temp, Clay Conc                                             &  66.68 \\
 SG Coating, Clay Conc, Dip Time Sol (s)                          &  81.03 \\
 Temp Alk Sol, Clay Conc, Dip Time Sol (s)                        &  81.03 \\
 NaOH Wt\%, Clay Conc, Dip Time Sol (s)                           &  81.03 \\
 Dip Time Alk (min), Clay Conc, Dip Time Sol (s)                  &  93.94 \\
 pH Sol, Clay Conc, Dip Time Sol (s)                              &  83.22 \\
 Hydro Time (h), Clay Conc, Dip Time Sol (s)                      &  56.92 \\
 Cure Time (min), Clay Conc, Dip Time Sol (s)                     &  60.09 \\
 Cure Temp, Clay Conc, Dip Time Sol (s)                           &  60.67 \\
 SG Coating, Clay Conc, Dip Time Sol (s), Dip Time Alk (min)      &  85.55 \\
 Temp Alk Sol, Clay Conc, Dip Time Sol (s), Dip Time Alk (min)    &  85.55 \\
 NaOH Wt\%, Clay Conc, Dip Time Sol (s), Dip Time Alk (min)       &  85.55 \\
 pH Sol, Clay Conc, Dip Time Sol (s), Dip Time Alk (min)          &  77.04 \\
 Hydro Time (h), Clay Conc, Dip Time Sol (s), Dip Time Alk (min)  &  75.85 \\
 Cure Time (min), Clay Conc, Dip Time Sol (s), Dip Time Alk (min) &  71.81 \\
 Cure Temp, Clay Conc, Dip Time Sol (s), Dip Time Alk (min)       &  63.33 \\
\hline
\end{tabular}
        \caption{Results of Forward Feature Selection for different features}
        \label{forward}
    \end{table*}