### 2024-08-09 NN/tables/backward.tex
zz

    \begin{table*}[h]
        \centering
        \begin{tabular}{lr}
\hline
 features                                                                                                                                 &    R2 \\
\hline
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp & 89.74 \\
 Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp             & 66.46 \\
 SG Coating, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp               & 66.46 \\
 SG Coating, Temp Alk Sol, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp            & 66.46 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp                     & 89.54 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp         &  9.16 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Dip Time Sol (s), Clay Conc, Cure Time (min), Cure Temp                 & 87.32 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Clay Conc, Cure Time (min), Cure Temp                   & 63.66 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Cure Time (min), Cure Temp            & 44.54 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Temp                  & 56.51 \\
 SG Coating, Temp Alk Sol, NaOH Wt\%, Dip Time Alk (min), pH Sol, Hydro Time (h), Dip Time Sol (s), Clay Conc, Cure Time (min)            & 66.27 \\
\hline
\end{tabular}
        \caption{Results of Backward Feature Elimination}
        \label{backward}
    \end{table*}