### 2024-08-09 NN/tables/forward.tex

    \begin{table*}[h]
        \centering
        \begin{tabular}{lr}
\hline
 features                                                            &     R2 \\
\hline
 pH Sol                                                              &  14.96 \\
 Hydro Time (h)                                                      &  33.65 \\
 Dip Time Sol (s)                                                    &  11.2  \\
 Clay Conc                                                           &  68.7  \\
 Cure Time (min)                                                     & -15.76 \\
 SG Coating, Clay Conc                                               &  63.12 \\
 Temp Alk Sol, Clay Conc                                             &  63.12 \\
 NaOH Wt\%, Clay Conc                                                &  63.12 \\
 Dip Time Alk (min), Clay Conc                                       &  47.64 \\
 pH Sol, Clay Conc                                                   &  77.65 \\
 Hydro Time (h), Clay Conc                                           & -24.51 \\
 Dip Time Sol (s), Clay Conc                                         &  68.92 \\
 Cure Time (min), Clay Conc                                          &  67.83 \\
 Cure Temp, Clay Conc                                                &  62.76 \\
 SG Coating, Clay Conc, pH Sol                                       &  82    \\
 Temp Alk Sol, Clay Conc, pH Sol                                     &  82    \\
 NaOH Wt\%, Clay Conc, pH Sol                                        &  82    \\
 Dip Time Alk (min), Clay Conc, pH Sol                               &  72.38 \\
 Hydro Time (h), Clay Conc, pH Sol                                   &  32.74 \\
 Dip Time Sol (s), Clay Conc, pH Sol                                 &  75.65 \\
 Cure Time (min), Clay Conc, pH Sol                                  &  70.52 \\
 Cure Temp, Clay Conc, pH Sol                                        &  71.84 \\
 Temp Alk Sol, Clay Conc, pH Sol, SG Coating                         &  87.46 \\
 NaOH Wt\%, Clay Conc, pH Sol, SG Coating                            &  87.46 \\
 Dip Time Alk (min), Clay Conc, pH Sol, SG Coating                   &  80.65 \\
 Hydro Time (h), Clay Conc, pH Sol, SG Coating                       &  43.08 \\
 Dip Time Sol (s), Clay Conc, pH Sol, SG Coating                     &  90.78 \\
 Cure Time (min), Clay Conc, pH Sol, SG Coating                      &  62.85 \\
 Cure Temp, Clay Conc, pH Sol, SG Coating                            &  58.38 \\
 Temp Alk Sol, Clay Conc, pH Sol, SG Coating, Dip Time Sol (s)       &  81.02 \\
 NaOH Wt\%, Clay Conc, pH Sol, SG Coating, Dip Time Sol (s)          &  81.02 \\
 Dip Time Alk (min), Clay Conc, pH Sol, SG Coating, Dip Time Sol (s) &  89.34 \\
 Hydro Time (h), Clay Conc, pH Sol, SG Coating, Dip Time Sol (s)     &  52.86 \\
 Cure Time (min), Clay Conc, pH Sol, SG Coating, Dip Time Sol (s)    &  73.46 \\
 Cure Temp, Clay Conc, pH Sol, SG Coating, Dip Time Sol (s)          &  67.5  \\
\hline
\end{tabular}
        \caption{Results of Forward Feature Selection for different features}
        \label{forward}
    \end{table*}