### 2024-08-29 NN/tables/backward.tex
commit

\begin{longtable}{lr}
\hline
 features                        &    R2 \\
\hline
 time, flowrate, temp, init conc & 98.54 \\
 flowrate, temp, init conc       &  0.66 \\
 time, temp, init conc           & 94.34 \\
 time, flowrate, init conc       & 98.02 \\
 time, flowrate, temp            & 95.41 \\
\hline
\end{longtable}