### 2024-08-29 NN/tables/forward.tex

\begin{longtable}{lr}
\hline
 features                        &    R2 \\
\hline
 time                            & 93.84 \\
 flowrate                        & -0.79 \\
 temp                            &  0.69 \\
 init conc                       &  1.47 \\
 flowrate, time                  & 95.32 \\
 temp, time                      & 93.29 \\
 init conc, time                 & 94.95 \\
 temp, time, flowrate            & 95.21 \\
 init conc, time, flowrate       & 97.84 \\
 temp, time, flowrate, init conc & 97.39 \\
\hline
\end{longtable}