### 2024-09-02 NN/plots_preds_latex.tex
commit

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots_preds/lstm-plots/R2-0.9655.pnt}
    \caption{R2-0.9655 - R-Squared: 0.97}
\end{figure}