### 2024-09-02 NN/plots_preds_latex.tex

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots_preds/R2-99.30-TanhReluFFNN-Impedance-module.png}
    \caption{Prediction of Impedance module with TanhReluFFNN epochs:200 - R-Squared: 0.99}
\end{figure}