### 2024-11-22 NN/plots_latex.tex
commit

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots/Linear1HiddenLayer-R2-vs-Num-of-Epochs-HiddenSize-10.png}
    \caption{Linear1HiddenLayer - R2 vs. Number of Epochs (Hidden Size: 10)}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots/Linear2HiddenLayer-R2-vs-Num-of-Epochs-HiddenSize-12.png}
    \caption{Linear2HiddenLayer - R2 vs. Number of Epochs (Hidden Size: 12)}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots/Linear2HiddenLayer-R2-vs-Num-of-Epochs-HiddenSize-20.png}
    \caption{Linear2HiddenLayer - R2 vs. Number of Epochs (Hidden Size: 20)}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots/Sig1HiddenLayer-R2-vs-Num-of-Epochs-HiddenSize-10.png}
    \caption{Sig1HiddenLayer - R2 vs. Number of Epochs (Hidden Size: 10)}
\end{figure}