### 2024-11-22 NN/plots_preds_latex.tex

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots_preds/R2-69.68-Sig1HiddenLayer-Temp.png}
    \caption{Prediction of Temp with Sig1HiddenLayer epochs:200 - R-Squared: 0.70}
\end{figure}