### 2024-11-22 NN/tables/forward.tex

\begin{longtable}{lr}
\hline
 features         &     R2 \\
\hline
 Pco2             &  -4.32 \\
 Pch4             & -14.19 \\
 rate             &  57.65 \\
 Pco2, rate       &  60.73 \\
 Pch4, rate       &  53.38 \\
 Pch4, rate, Pco2 &  41.27 \\
\hline
\end{longtable}