### 2024-11-22 NN/plots_preds_latex.tex
commit

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{plots_preds/R2-91.77-Tanh2HiddenLayer-rate.png}
    \caption{Prediction of rate with Tanh2HiddenLayer epochs:50 - R-Squared: 0.92}
\end{figure}